---
title: "Data Exploration, ITT, CACE, and CACE with IVY"
output: pdf_document
---

```{r}
library('data.table')
library('readr')
library('ggplot2')
library('readxl')
library('tidyverse')
library('sandwich')
library('lmtest')
library("ivreg")
```
# w241 Final Project: Clean Data and Basic Analysis

## Read in data

```{r}
# read in data files
treatment_initial <- read_excel('groupA_initial.xlsx')
control_initial <- read_excel("groupB_initial.xlsx")
treatment_final <- read_excel('groupA_final.xlsx')
control_final <- read_excel("groupB_final.xlsx")

```



## Cleaning data
```{r}
# dropping and renaming columns

clean_initial_data <- function(data, treatment = FALSE) {
  
  # dropping first, extra row
  names(data) <- as.matrix(data[1, ])
  data <- data[-1, ]
  data[] <- lapply(data, function(x) type.convert(as.character(x)))
  
  # dropping and renaming columns for CONTROL GROUP
  if (treatment == FALSE) {
    data <- data[,-c(1:2)]
    data <- data[,-c(10:19)]
    data <- data[,-c(11)]
    data <- data.table(data)
    new <- c('date', 'first_name', 'last_name', 'age', 'country', 'social_apps', 'iphone', 'screen_time_rep', 'screen_time_on', 'screen_time_avg_initial')
    setnames(data,colnames(data),new)}
  
  # dropping and renaming columns for TREATMENT GROUP
  else {
    # dropping columns
    data <- data[,-c(1:7)]
    data<- data[,-c(2:10)]
    data <- data[,-c(10:19)]
    data <- data[,-c(12)]
    data <- data.table(data)
    new <- c('date', 'first_name', 'last_name', 'age', 'country', 'social_apps', 'iphone', 'greyscale', 'screen_time_rep', 'screen_time_on', 'screen_time_avg_initial')
    setnames(data,colnames(data),new)
  }
  
  # get rid of rows with empty values for name (indicates a faulty response)
  complete_vec <- complete.cases(data[, 'first_name'])
  data <- data[complete_vec,]
  
  # standardizing 'United States' label
  data$country <- ifelse((data$country == 'Mexico')|(data$country == 'MX')|(data$country == 'Mx')|(data$country == 'mexico')|(data$country == 'México'),             'Mexico', 'United States' )
  # convert date to readable type
  data$date <- as.Date(data$date, origin = "1899-12-30")
  
  # create a variable to indicate which group (first cohort or second) based on date
  data[,first_group := (as.numeric(as.Date(data$date, origin = "1899-12-30") < "2021-03-23"))]
  
  return(data)
}
```

```{r}
# dropping and renaming columns 

clean_final_data <- function(data, treatment = FALSE) {
  
  # dropping first, extra row
  names(data) <- as.matrix(data[1, ])
  data <- data[-1, ]
  data[] <- lapply(data, function(x) type.convert(as.character(x)))
  
  # dropping columns
  data <- data[,-c(1:7)]
  data <- data[,-c(2:10)]
  data <-data[,-c(5:28)]
  
  # dropping and renaming columns for CONTROL GROUP
  if (treatment == FALSE) {
    data <- data[,-c(8)]
    data <- data.table(data)
    new <- c('date', 'first_name', 'last_name', 'iphone', 'screen_time_1', 'screen_time_2', 'screen_time_3')
    setnames(data,colnames(data),new)}
  
  # dropping and renaming columns for TREATMENT GROUP
  else {
    # dropping columns
    data <- data.table(data)
    new <- c('date', 'first_name', 'last_name', 'iphone', 'screen_time_1', 'screen_time_2', 'screen_time_3')
    setnames(data,colnames(data),new)
  }
  
  # get rid of rows with empty values for name (indicates a faulty response)
  complete_vec <- complete.cases(data[, 'first_name'])
  data <- data[complete_vec,]
  
  # averages screen time 
  data[, screen_time_avg_final := (rowMeans(data[,c('screen_time_1', 'screen_time_2', 'screen_time_3')]))]
  
  # convert date to readable type
  data$date <- as.Date(data$date, origin = "1899-12-30")
  
  # create a variable to indicate which group (first cohort or second) based on date
  data[,first_group := (as.numeric(as.Date(data$date, origin = "1899-12-30") < "2021-03-26"))]
  
  return(data)
}
```


```{r}
treatment_initial <- clean_initial_data(treatment_initial, treatment = TRUE)
control_initial <- clean_initial_data(control_initial, treatment = FALSE)
treatment_final <- clean_final_data(treatment_final, treatment = TRUE)
control_final <- clean_final_data(control_final, treatment = FALSE)
```


```{r}
# getting rid of rows in final that don't exist in the initial (calling these faulty responses)
# joining initial and final datasets for both treatment and control groups

treatment_final <- treatment_final[ treatment_final$first_name %in% treatment_initial$first_name, ]
control_final <- control_final[ control_final$first_name %in% control_initial$first_name, ]
treatment <- merge(treatment_initial, treatment_final[,c('first_name','screen_time_avg_final')], by.x = "first_name", by.y = "first_name", all.x = TRUE, all.y = FALSE)
control <- merge(control_initial, control_final[,c('first_name','screen_time_avg_final')], by.x = "first_name", by.y = "first_name", all.x = TRUE, all.y = FALSE)

```


```{r}
# replacing NA with zero for groups
treatment[treatment == 0] <- NA
control[control == 0] <- NA

# creating treatment column
treatment$treatment <- 1
control$treatment <- 0

# binding treatment and control data into one dataframe
data <- rbind(treatment,control, fill=TRUE)

# drop people who have NA values for both screen_time_avg_final and screen_time_avg_initial
# issue with qualtrics for not getting their data for some reason
data <- data[!with(data,is.na(screen_time_avg_initial)& is.na(screen_time_avg_final)),]
```


## Plot histograms of cleaned data

```{r}
ggplot(data, aes(`screen_time_avg_initial`, fill = factor(treatment))) + geom_histogram(alpha = 0.2, position = 'dodge') + labs(title="Baseline Screen Time Averages")
```

```{r}
ggplot(data, aes(`screen_time_avg_final`, fill = factor(treatment))) + geom_histogram(alpha = 0.2, position = 'dodge') + labs(title="Final Screen Time Averages")
```


```{r}
# replace na values in baseline (screen_time_avg_initial) with average of that column
# this is okay because we want an average baseline of where people are
# we don't want to do this with screen_time_avg_final because of attrition (we want to know which people didn't come back)
# we do this separately for treatment and control group
data$screen_time_avg_initial[is.na(data$screen_time_avg_initial) & data$treatment == 1] <- mean(treatment$screen_time_avg_initial,na.rm=TRUE)

data$screen_time_avg_initial[is.na(data$screen_time_avg_initial) & data$treatment == 0] <- mean(control$screen_time_avg_initial,na.rm=TRUE)

# getting the difference between final and initial screen times
data[, diff := screen_time_avg_final - screen_time_avg_initial]
```


```{r}
# compare individuals' initial and final averages
ggplot(data, aes(`diff`, fill = factor(treatment))) + geom_histogram(alpha = 0.2, position = 'dodge') + labs(title="Screen Time Average Difference")
```


## Estimating ATE, ITT, and CACE ignoring attrition

```{r}
# making greyscale (receiving treatment) and iphone variables binary for analysis

data[, greyscale_binary := ifelse(greyscale =='Yes', 1, 0)]
data[is.na(greyscale_binary), greyscale_binary:=0]
# makes iphone column binary (iphone = 1, android = 0)
data$iphone <- as.numeric(data$iphone == 'IPhone')
```

```{r}
# estimating average treatment effect
ate <- data[, lm(screen_time_avg_final~ treatment + screen_time_avg_initial + as.factor(country) + iphone)]
ate$vcovHC_ <- vcovHC(ate)
coeftest(ate, vcov. = ate$vcovHC_)
```

This analysis regresses final average screen time (`screen_time_avg_final`) on treatment assignment (`treatment`), while controlling for initial average screen time (`screen_time_avg_initial`), country, and phone. Results suggest that the average treatment effect is -0.00033505 with a 95% confidence interval of +/- 1.272597 and that there is not a statistically significant treatment effect on final average screen time.


```{r}
# estimating the ITT

itt <- data[, lm(screen_time_avg_final ~ treatment)]
itt$vcovHC_ <- vcovHC(itt)
coeftest(itt, vcov. = itt$vcovHC_)

```

In this analysis, the final average screen time (`screen_time_avg_final`) is regressed on treatment assignment (`treatment`). Screen time is a continuous variable in units of hours and treatment assignment (`treatment`) is scored 1 if the subject was assigned to the treatment group and 0 otherwise. The ITT is 0.042928 with a 95% confidence interval of 1.730606. This result is not statistically significant.

```{r}
# using ivyreg to calculate CACE
ivyreg <- data[, ivreg(screen_time_avg_final ~ greyscale_binary, ~treatment)]
ivyreg$vcovHC_ <- vcovHC(ivyreg)
coeftest(ivyreg, vcov. = ivyreg$vcovHC_)

```

In this instrumental variables regression model, the final average screen time (`screen_time_avg_final`) is regressed on actual treatment (`greyscale_binary`) using treatment assignment(`treatment`) as an instrument. The results suggest that turning on greyscale increased screentime usage among Compliers by 4.7% with a 95% confidence interval of 1.89498. However, the results suggest that this effect is not statistically insignificant. 


## Extreme Error Bounds

### Minimum bound

```{r}
#  assume that all of the data that you did not observe was actually a very small value
min <- data[, min(screen_time_avg_final, na.rm = TRUE)]
data[, low_screen_time := screen_time_avg_final]
data[is.na(low_screen_time), low_screen_time := min]
```

```{r}
# estimating average treatment effect
ate_min <- data[, lm(low_screen_time ~ treatment + screen_time_avg_initial + as.factor(country) + iphone)]
ate_min$vcovHC_ <- vcovHC(ate_min)
coeftest(ate_min, vcov. = ate_min$vcovHC_)
```

This analysis regresses final average screen time, bounded by the extreme minimum,  (`low_screen_time`) on treatment assignment (`treatment`), while controlling for initial average screen time (`screen_time_avg_initial`), country, and phone. Results suggest that the treatment effect is 0.95687 with a 95% confidence interval of 1.23602 and that there is not a statistically significant treatment effect on final average screen time.

```{r}
# estimating the ITT

itt_min <- data[, lm(low_screen_time ~ treatment)]
itt_min$vcovHC_ <- vcovHC(itt_min)
coeftest(itt_min, vcov. = itt_min$vcovHC_)
```

In this analysis, the final average screen time, bounded by the extreme minimum, (`low_screen_time`) is regressed on treatment assignment (`treatment`). Screen time is a continuous variable in units of hours and treatment assignment (`treatment`) is scored 1 if the subject was assigned to the treatment group and 0 otherwise. The ITT is 1.03392 with a 95% confidence interval of 1.37848. This result is not statistically significant.


```{r}
# using ivyreg to calculate CACE
ivyreg_min <- data[, ivreg(low_screen_time ~ greyscale_binary, ~treatment)]
ivyreg_min$vcovHC_ <- vcovHC(ivyreg_min)
coeftest(ivyreg_min, vcov. = ivyreg_min$vcovHC_)
```

In this instrumental variables regression model, the final average screen time, bounded by the extreme minimum, (`low_screen_time`) is regressed on actual treatment (`greyscale_binary`) using treatment assignment(`treatment`) as an instrument. The results suggest that turning on greyscale increased screentime usage among Compliers by 115.32% with a 95% confidence interval of 1.52946. However, the results suggest that this effect is not statistically insignificant.

### Maximum bound

```{r}
#  assume that all of the data that you did not observe was actually a very high value
max <- data[, max(screen_time_avg_final, na.rm = TRUE)]
data[, high_screen_time := screen_time_avg_final]
data[is.na(high_screen_time), high_screen_time := max]
```

```{r}
# estimating average treatment effect
ate_max <- data[, lm(high_screen_time ~ treatment + screen_time_avg_initial + as.factor(country) + iphone)]
ate_max$vcovHC_ <- vcovHC(ate_max)
coeftest(ate_max, vcov. = ate_max$vcovHC_)
```

This analysis regresses final average screen time, bounded by the extreme maximum,  (`high_screen_time`) on treatment assignment (`treatment`), while controlling for initial average screen time (`screen_time_avg_initial`), country, and phone. Results suggest that the treatment effect is -1.58117 with a 95% confidence interval of 1.85502 and that there is a slightly statistically significant treatment effect on final average screen time.

```{r}
# estimating the ITT

itt_max <- data[, lm(high_screen_time ~ treatment)]
itt_max$vcovHC_ <- vcovHC(itt_max)
coeftest(itt_max, vcov. = itt_max$vcovHC_)
```

In this analysis, the final average screen time, bounded by the extreme maximum, (`high_screen_time`) is regressed on treatment assignment (`treatment`). Screen time is a continuous variable in units of hours and treatment assignment (`treatment`) is scored 1 if the subject was assigned to the treatment group and 0 otherwise. The ITT is -1.57130 with a 95% confidence interval of 1.75706. This result is slightly statistically significant.


```{r}
# using ivyreg to calculate CACE
ivyreg_max <- data[, ivreg(high_screen_time ~ greyscale_binary, ~treatment)]
ivyreg_max$vcovHC_ <- vcovHC(ivyreg_max)
coeftest(ivyreg_max, vcov. = ivyreg_max$vcovHC_)
```

In this instrumental variables regression model, the final average screen time, bounded by the extreme maximum, (`high_screen_time`) is regressed on actual treatment (`greyscale_binary`) using treatment assignment(`treatment`) as an instrument. The results suggest that turning on greyscale increased screentime usage among Compliers by -175.26% with a 95% confidence interval of 1.9713. However, the results suggest that this effect is slightly statistically insignificant.

### Mean bound

```{r}
#  assume that all of the data that you did not observe was actually the average value
mean <- data[ , mean(screen_time_avg_final, na.rm = TRUE)]

data[ , mean_screen_time := screen_time_avg_final]
data[is.na(mean_screen_time), mean_screen_time := mean]
ate_mean <- data[, lm(mean_screen_time ~ treatment + screen_time_avg_initial + as.factor(country) + iphone)]
ate_mean$vcovHC_ <- vcovHC(ate_mean)
coeftest(ate_mean, vcov. = ate_mean$vcovHC_)
```

This analysis regresses final average screen time, bounded by the mean,  (`mean_screen_time`) on treatment assignment (`treatment`), while controlling for initial average screen time (`screen_time_avg_initial`), country, and phone. Results suggest that the treatment effect is -0.024055 with a 95% confidence interval of 0.884826 and that there is not a statistically significant treatment effect on final average screen time.

```{r}
# estimating the ITT

itt_mean <- data[, lm(mean_screen_time ~ treatment)]
itt_mean$vcovHC_ <- vcovHC(itt_mean)
coeftest(itt_mean, vcov. = itt_mean$vcovHC_)
```

In this analysis, the final average screen time, bounded by the mean, (`mean_screen_time`) is regressed on treatment assignment (`treatment`). Screen time is a continuous variable in units of hours and treatment assignment (`treatment`) is scored 1 if the subject was assigned to the treatment group and 0 otherwise. The ITT is 0.027034 with a 95% confidence interval of 1.061992. This result is not statistically significant.


```{r}
# using ivyreg to calculate CACE
ivyreg_mean <- data[, ivreg(mean_screen_time ~ greyscale_binary, ~treatment)]
ivyreg_mean$vcovHC_ <- vcovHC(ivyreg_mean)
coeftest(ivyreg_mean, vcov. = ivyreg_mean$vcovHC_)
```

In this instrumental variables regression model, the final average screen time, bounded by the mean, (`mean_screen_time`) is regressed on actual treatment (`greyscale_binary`) using treatment assignment(`treatment`) as an instrument. The results suggest that turning on greyscale increased screentime usage among Compliers by 3.0153% with a 95% confidence interval of 1.184236. However, the results suggest that this effect is not statistically insignificant.

## Table of extreme value bands
```{r}
stargazer(
  ate, 
  ate_min, 
  ate_max,
  ate_mean,
  type = 'text',
  se = list(
    robust_se(ate), 
    robust_se(ate_min),
    robust_se(ate_max),
    robust_se(ate_mean)
  ), 
  omit.stat = c('ser', 'F')
)
```